### 2020-04-04 Thesis/conc.tex
Corrections.
Formatting/spell changes

Let us try this out with an example.  
Fix $X\in\{\bB,\bN\}$ and let \emph{Nondeterministic PCF} be the language PCF, together with a constant $\ask_X \from \bool$ that plays the role of a nondeterministic oracle.  
It is the language with types given by
\[
  T \Coloneqq \bool \mid \nat \mid T \to T\,,
  \]
and with a type theory as shown in Figure \ref{FigNDPcfTypeTheory}.
\begin{figure}
  \begin{mathpar}
    \inferrule*{ }{\Gamma,x\from T \ts x \from T}
    \and
    \inferrule*{\Gamma\ts M \from S \to T \\ \Gamma \ts N \from S}{\Gamma\ts MN \from T}
    \and
    \inferrule*{\Gamma,x\from S \ts M \from T}{\Gamma \ts \lambda x^S.M \from S \to T}
    \and
    \inferrule*{ }{\Gamma\ts\true\from\bool}
    \and
    \inferrule*{ }{\Gamma\ts\false\from\bool}
    \\
    \inferrule*[right={$T\in\{\bool,\nat\}$}]{\Gamma\ts M \from \bool \\ \Gamma \ts N \from T \\ \Gamma \ts P \from T}{\Gamma\ts \If M \Then N \Else P \from T}
    \\
    \inferrule*{ }{\Gamma\ts n\from\nat}
    \and
    \inferrule*{\Gamma\ts M \from \nat}{\Gamma\ts \suc M \from \nat}
    \and
    \inferrule*{\Gamma\ts M \from \nat}{\Gamma\ts \pred M\from \nat}
    \\
    \inferrule*[right={$T\in\{\bool,\nat\}$}]{\Gamma\ts M \from \nat \\ \Gamma\ts N \from T \\ \Gamma \ts P \from T}{\Gamma \ts \IfO M \Then N \Else P \from T}
    \and
    \inferrule*{\Gamma \ts M \from T \to T}{\Gamma \ts \Y_T M \from T}
    \and
    \inferrule*{ }
    {\ask_X \from\bool}
  \end{mathpar}
  \caption[Type theory for a variant of PCF with nondeterminism.]
  {Type theory for a variant of PCF with nondeterminism.
  Note that if the nondeterminism is infinite (i.e., $X=\bN$), then it is common (see \cite{LairdOrdinalGames} for nondeterminism and \cite{ProbabilisicPcf} for the probabilistic case) to modify the conditional construct $\IfO M\Then N \Else P$ so that the value of $M$ is passed into $P$ (so that the typing rule for $\IfO$ becomes
  \[
    \inferrule*[right={$T\in\{\bool,\nat\}$}]{\Gamma\ts M \from \nat \\ \Gamma \ts N \from T \\ \Gamma \ts P \from \nat \to T}
    {\Gamma \ts \IfO M \Then N \Else P \from T}\,.
    \]
  The idea behind this is that if $M$ evaluates to $0$ then $N$ will be evaluated, and if it evaluates to $n+1$ then $P\,n$ will be evaluated.
  In ordinary deterministic PCF, this does not change the expressive power of the language, since we could always write
  \[
    \IfO M \Then N \Else (P\,M)\,.
    \]
  However, this does not work if $M$ may exhibit nondeterministic behaviour, since $M$ may evaluate to two different values the two times it is called.
  We will stick with the variant shown above for the sake of simplicity, but nothing changes in our proofs if we choose the other version.}
  \label{FigNDPcfTypeTheory}
\end{figure}
We endow this language with an operational semantics of may testing.  
Define a \emph{canonical form} of the language to be a term taking one of the following forms.
\begin{itemize}
  \item At type $\bool$, the constants $\true$ and $\false$;
  \item at type $\nat$, the numerals $n$; and
  \item at type $S \to T$, terms of the form $\lambda x^S.M$.
\end{itemize}
We then define a relation $M \converges c$ (read `$M$ \emph{may converge to} $c$'), for $M$ a term of Nondeterministic PCF and $c$ a canonical form, inductively as in Figure \ref{FigNDPCFOpSem}.
This rule is the restriction of the $\converges$ rule from Nondeterministic Idealized Algol to the set of those terms that are terms of Nondeterministic PCF.
\begin{figure}
  \begin{mathpar}
    \inferrule*{ }{c\converges c}
    \and
    \inferrule*{M\converges\lambda x.M' \\ M'[N/x] \converges c}{M\,N\converges c}
    \\
    \inferrule*{M\converges \true \\ N \converges c}{\If M \Then N \Else P \converges c}
    \and
    \inferrule*{M\converges \false \\ P \converges c}{\If M \Then N \Else P \converges c}
    \\
    \inferrule*{M \converges n}{\suc M \converges n+1}
    \and
    \inferrule*{M \converges n+1}{\pred M \converges n}
    \and
    \inferrule*{M \converges 0}{\pred M \converges 0}
    \\
    \inferrule*{M \converges 0 \\ N \converges c}{\IfO M \Then N \Else P \converges c}
    \and
    \inferrule*{M \converges n+1 \\ P \converges c}{\IfO M \Then N \Else P \converges c}
    \\
    \inferrule*{M(\Y M) \converges c}{\Y M \converges c}
    \and
    \inferrule*{ }{\ask_X \converges \true}
    \and
    \inferrule*{ }{\ask_X \converges \false}
  \end{mathpar}
  \caption{Big-step operational semantics for Nondeterministic PCF and may testing}
  \label{FigNDPCFOpSem}
\end{figure}

>>>>

Let us try this out with an example.  
Fix $X\in\{\bB,\bN\}$ and let \emph{Nondeterministic PCF} be the language PCF, together with a constant $\ask_X \from \bool$ that plays the role of a nondeterministic oracle.  
It is the language with types given by
\[
  T \Coloneqq \bool \mid \nat \mid T \to T\,,
  \]
and with a type theory as shown in Figure \ref{FigNDPcfTypeTheory}.
\begin{figure}
  \begin{mathpar}
    \inferrule*{ }{\Gamma,x\from T \ts x \from T}
    \and
    \inferrule*{\Gamma\ts M \from S \to T \\ \Gamma \ts N \from S}{\Gamma\ts MN \from T}
    \and
    \inferrule*{\Gamma,x\from S \ts M \from T}{\Gamma \ts \lambda x^S.M \from S \to T}
    \and
    \inferrule*{ }{\Gamma\ts\true\from\bool}
    \and
    \inferrule*{ }{\Gamma\ts\false\from\bool}
    \\
    \inferrule*[right={$T\in\{\bool,\nat\}$}]{\Gamma\ts M \from \bool \\ \Gamma \ts N \from T \\ \Gamma \ts P \from T}{\Gamma\ts \If M \Then N \Else P \from T}
    \\
    \inferrule*{ }{\Gamma\ts n\from\nat}
    \and
    \inferrule*{\Gamma\ts M \from \nat}{\Gamma\ts \suc M \from \nat}
    \and
    \inferrule*{\Gamma\ts M \from \nat}{\Gamma\ts \pred M\from \nat}
    \\
    \inferrule*[right={$T\in\{\bool,\nat\}$}]{\Gamma\ts M \from \nat \\ \Gamma\ts N \from T \\ \Gamma \ts P \from T}{\Gamma \ts \IfO M \Then N \Else P \from T}
    \and
    \inferrule*[right={$S,T\in\{\bool,\nat\}$}]{\Gamma\ts M \from S \\ \Gamma,x\from S\ts N \from T}{\Gamma\ts \lett x=M\inn N\from T}
    \and
    \inferrule*{\Gamma \ts M \from T \to T}{\Gamma \ts \Y_T M \from T}
    \and
    \inferrule*{ }
    {\ask_X \from\bool}
  \end{mathpar}
  \caption[Type theory for a variant of PCF with nondeterminism.]
  {Type theory for a variant of PCF with nondeterminism.
  Note that although ordinary deterministic PCF does not include the $\lett$ construct, it has to be included in the nondeterministic (see \cite{LairdOrdinalGames}) and probabilistic (see \cite{ProbabilisicPcf}) versions.

  In ordinary PCF, we may simulate the term $\lett x=M \inn N$ by the term $(\lambda x.N) M$, which is less efficient than the version with $\lett$ (If $N$ contains multiple occurrences of the variable $x$, then we end up compute $M$ multiple times), but which still gives the same result.

  In nondeterministic variants of PCF, however, evaluating $M$ multiple times may give a different answer each time, in contrast to the version with $\lett$, in which we get a single answer and use it multiple times.}
  \label{FigNDPcfTypeTheory}
\end{figure}
We endow this language with an operational semantics of may testing.  
Define a \emph{canonical form} of the language to be a term taking one of the following forms.
\begin{itemize}
  \item At type $\bool$, the constants $\true$ and $\false$;
  \item at type $\nat$, the numerals $n$; and
  \item at type $S \to T$, terms of the form $\lambda x^S.M$.
\end{itemize}
We then define a relation $M \converges c$ (read `$M$ \emph{may converge to} $c$'), for $M$ a term of Nondeterministic PCF and $c$ a canonical form, inductively as in Figure \ref{FigNDPCFOpSem}.
This rule is the restriction of the $\converges$ rule from Nondeterministic Idealized Algol to the set of those terms that are terms of Nondeterministic PCF.
\begin{figure}
  \begin{mathpar}
    \inferrule*{ }{c\converges c}
    \and
    \inferrule*{M\converges\lambda x.M' \\ M'[N/x] \converges c}{M\,N\converges c}
    \\
    \inferrule*{M\converges \true \\ N \converges c}{\If M \Then N \Else P \converges c}
    \and
    \inferrule*{M\converges \false \\ P \converges c}{\If M \Then N \Else P \converges c}
    \\
    \inferrule*{M \converges n}{\suc M \converges n+1}
    \and
    \inferrule*{M \converges n+1}{\pred M \converges n}
    \and
    \inferrule*{M \converges 0}{\pred M \converges 0}
    \\
    \inferrule*{M \converges 0 \\ N \converges c}{\IfO M \Then N \Else P \converges c}
    \and
    \inferrule*{M \converges n+1 \\ P \converges c}{\IfO M \Then N \Else P \converges c}
    \\
    \inferrule*{M(\Y M) \converges c}{\Y M \converges c}
    \and
    \inferrule*{ }{\ask_X \converges \true}
    \and
    \inferrule*{ }{\ask_X \converges \false}
  \end{mathpar}
  \caption{Big-step operational semantics for Nondeterministic PCF and may testing}
  \label{FigNDPCFOpSem}
\end{figure}
